---
author: "Thomas Lamot, Laloux lab"
output: html_notebook
title: "Pipeline for growth & cell cycle analysis of Bdellovibrio bacteriovorus"
---

# Dependencies

This pipeline has the following dependencies: *yaml, here, furrr, runner, sf, tidyverse, terra, fs, glue, uuid, splus2R, janitor*.

# Global settings: locate the parameter file.

Change the path to *setting_file* to locate your 'setting.yml'.

```{r}
here::i_am("Pipeline/Pipeline_final.Rmd")
library(here)
here()
setting_file <- "E:/Yoann/M9_ParB/setting.yml"
```


# Load libraries and setting

The following chunk loads the required libraries and locates the microscopy images.

```{r}
library(tidyverse)
library(sf)
library(terra)
library(furrr)


source(here("Pipeline", "extract_from.R"), encoding = "UTF-8")
source(here("Pipeline", "readers.R"), encoding = "UTF-8")
source(here("Pipeline", "helperfunctions.R"), encoding = "UTF-8")


settings <-  yaml::read_yaml(setting_file)

meshes_glob <- switch (settings$Analysis_settings$Meshes$Type,
  "Oufti" = "*.mat",
  "MicrobeJ" = "*.csv",
  "R" = "*.rds"
)

meshes_fct <- switch (settings$Analysis_settings$Meshes$Type,
  "Oufti" = read_oufti,
  "MicrobeJ" = read_microbeJ_mesh,
  "R" = readRDS
)

predator_pix_fct <- switch (settings$Analysis_settings$Diffuse$Predator_pixels_threshold$Type,
  "median_mad" = median_mad_threshold,
  "mean_sd" = mean_sd_threshold
)

central_size_fct <- switch (settings$Analysis_settings$Diffuse$Growth$Central_size_threshold_type,
  "mean" = mean,
  "median" = median)


phase_images   <- NULL
focis_images   <- NULL
diffuse_images <- NULL

if (!is.null(settings$Analysis_settings$Folders$Phase_folder)) {
  phase_images <- fs::dir_ls(settings$Analysis_settings$Folders$Phase_folder, glob ="*.tif*", recurse = TRUE)
}
if (!is.null(settings$Analysis_settings$Folders$Foci_folder)) {
  focis_images <- fs::dir_ls(settings$Analysis_settings$Folders$Foci_folder, glob ="*.tif*", recurse = TRUE)
}
if (!is.null(settings$Analysis_settings$Folders$Diffuse_folder)) {
  diffuse_images <- fs::dir_ls(settings$Analysis_settings$Folders$Diffuse_folder, glob ="*.tif*", recurse = TRUE)
}

if (is.null(phase_images)) {
  print("Warning: No phase images detected")
  }
if (length(phase_images) != length(focis_images) & !is.null(focis_images)) {
  print("Warning: Number of images in phase and focis do not match")
  print(setdiff(phase_images, focis_images))
  }
if (length(phase_images) != length(diffuse_images) & !is.null(diffuse_images)) {
  print("Warning: Number of images in phase and diffuse do not match")
  print(setdiff(phase_images, diffuse_images))
}

```


# Mesh loading

The following chunk loads the input of the cell outlines derived from Oufti or MicrobeJ (referred to as 'meshes' in this pipeline), and computes the buffered outline & prey cell dimensions.

```{r}

# Needed because when using oufti in batch mode, the output file only save the order of the analysed images
images_infos <- 
  # Correspondence between the splitted phase, diffuse and foci images & info on which condition and replicate they belong 
  # Condition folder name should be the same as the corresponding oufti result file
  tibble(
    phase_image_path   = phase_images,
    phase_image_name   = phase_images |> fs::path_file(),
    focis_image_path   = if(is.null(focis_images))   {NA_character_} else {focis_images},
    focis_image_name   = if(is.null(focis_images))   {NA_character_} else {focis_images   |> fs::path_file()},
    diffuse_image_path = if(is.null(diffuse_images)) {NA_character_} else {diffuse_images},
    diffuse_image_name = if(is.null(diffuse_images)) {NA_character_} else {diffuse_images |> fs::path_file()},
    replicate = phase_images |> fs::path_dir() |> fs::path_file(),
    condition = phase_images |> fs::path_dir() |> fs::path_dir() |> fs::path_file()) |>
  arrange(phase_image_name) |>
  rowwise() |>
  mutate(
    # Alphabetical order of the images -> should be the same of the analysed images in oufti
    id = paste(condition, replicate, uuid::UUIDgenerate(), sep = "-")) |>
  group_by(condition) |>
  mutate(
    image_id = row_number()) |>
  # Discard image analysed present in the oufti result file, avoid redoing the analysis while preserving the order
  filter(!(replicate %in% settings$Analysis_settings$Folders$Bad_replicate_name))

# oufti result files
mesh_files <- fs::dir_ls(settings$Analysis_settings$Folders$Meshes_folder, glob = meshes_glob) 

buffer_pix <- compute_buffer_pix(settings$Analysis_settings$Meshes$Buffer_um, settings$Analysis_settings$Microscopy_info$Pixel_to_um_ratio)

meshes <- map_df(mesh_files, \(x) meshes_fct(x), .id = "filename") |>
    select(-c("algorithm", "birthframe", "polarity", "ancestors", "descendants", "stage", "timelapse", "divisions", "box")) |>
    mutate(condition = fs::path_file(filename) |> fs::path_ext_remove()) |>
    st_sf() |>
    rename(image_id = frame) |>
    inner_join(images_infos, by = c("condition", "image_id")) |>
    rowwise() |>
    mutate(
      mesh_pix = mesh,
      mesh     = mesh * settings$Analysis_settings$Microscopy_info$Pixel_to_um_ratio,
#      max.length = max.length * settings$Analysis_settings$Microscopy_info$Pixel_to_um_ratio,
#      max.width  = max.width  * settings$Analysis_settings$Microscopy_info$Pixel_to_um_ratio,
      bdelloplast_area   = st_area(mesh),
#      cent_x = st_centroid(mesh)[[1]][[1]],
#      cent_y = st_centroid(mesh)[[1]][[2]],
      cell   = as.character(cell),
#      radius = mean(st_distance(st_cast(mesh, "MULTIPOINT"), st_centroid(mesh))),
#      estimated_area   = circle_area(radius),
#      estimated_volume = sphere_volume(radius),
      id = paste(id, cell, sep = "-"),
      config_file = fs::path_file(setting_file)
    ) |>
    group_by(phase_image_name) |>
    mutate(mesh_buffered = st_make_valid(sf::st_buffer(mesh_pix, dist = buffer_pix, nQuadSegs = 5)))


rm(mesh_files, images_infos)
gc()

```

# Popping points computation phase

In phase contrast moviestime-lapses, there is a sharp decrease in contrast between the bdelloplast and its surroundings when Bdellovibrio divides. This was used as a proxy to determine the popping time. The contrast was computed as the weighted standard deviation of the phase contrast pixels within the buffered bdelloplast outline. Here each pixel was weighed by its coverage fraction within the bdelloplast outline. Finally, the maximum decrease in contrast is computed as the maximum difference of two successive running means of the contrast values. 

```{r}
if (!is.null(settings$Analysis_settings$Folders$Phase_folder)) {

l_meshes  <- get_mesh_list(meshes, mesh_buffered, id, phase_image_path)
img_paths <- get_image_list(meshes, phase_image_path)

phase_intensity <- extract_image_pixels_from_meshes(img_paths, l_meshes, include_cols = c("id"), parallel = TRUE) 

rm(l_meshes)
gc()

phase_intensity |>
  group_by(id, frame)|>
  summarise(
    wmean    = weighted.mean(value, coverage_fraction, na.rm = TRUE),
    contrast = weigthed_sd(value, coverage_fraction)) |> 
  group_by(id) |>
  arrange(frame) |>
  mutate(
    delta_mean_contrast = running_means_difference(
      contrast,
      span = settings$Analysis_settings$Popping_point_phase$Running_means_span),
    is_local_peak = splus2R::peaks(
      delta_mean_contrast,
      span = settings$Analysis_settings$Popping_point_phase$Peaks_finding_span,
      strict = FALSE,
      endbehavior = 1),
    has_peak = TRUE %in% is_local_peak) -> phase_intensity_summary 
  
rm(phase_intensity)
gc()

if (settings$Analysis_settings$Popping_point_phase$Save_intermediate_results) {
  if (exists("phase_intensity_summary")) {
    saveRDS(phase_intensity_summary, fs::path(settings$Analysis_settings$Folders$Analysis_folder) / "phase_intensity_summary.rds")
  }
}


popping_frames <- phase_intensity_summary |>
  dplyr::filter(has_peak) |>
  mutate(
     popping_frame = find_min(frame, for_which = is_local_peak & frame >= settings$Analysis_settings$Popping_point_phase$Minimum_frame)
    ) |> 
  filter(frame == popping_frame) |>
  select(id, popping_frame)


  
rm(phase_intensity_summary)
gc()

}
```

# Focis detection

Foci are detected by TrackMate, a standard plugin in the Fiji distribution of ImageJ. Make sure you have this installed: https://imagej.net/software/fiji/

```{r}
if (!is.null(settings$Analysis_settings$Folders$Foci_folder)) {

for_detection <- meshes |> 
  as_tibble() |>
  select(focis_image_path, focis_image_name, condition, replicate) |>
  distinct() |>
  mutate(foci_out_path = fs::path(settings$Analysis_settings$Folders$Foci_detection_folder) / condition / replicate / fs::path_ext_set(focis_image_name, ".csv"))

collapser <- "', '"
focis_out_py <- glue::glue("['{paste(for_detection$foci_out_path, collapse = collapser)}']")
focis_image_py <- glue::glue("['{paste(for_detection$focis_image_path, collapse = collapser)}']")

fs::dir_create(for_detection$foci_out_path |>fs::path_dir(), recurse = TRUE)

macro <- glue::glue(read_file(here("Pipeline", "TrackMate_batch_Jython_template.txt")))
    
tmp_macro_file <- fs::file_temp(tmp_dir = fs::path(fs::path_dir(settings$Global_settings$fiji_path)) / "macros/", ext = "py")
cat(macro, file = tmp_macro_file)
command <- glue::glue("{settings$Global_settings$fiji_path} --run {tmp_macro_file}")
system(command, wait = TRUE, ignore.stdout = FALSE)
    
fs::file_delete(tmp_macro_file)

gc()

}
```


# Focis analysis

This chunk counts the number of foci in each bdelloplast.

```{r}
if (!is.null(settings$Analysis_settings$Folders$Foci_folder)) {

# replicates and conditions not necessary since we have the ids
detected_foci_files <- fs::dir_ls(settings$Analysis_settings$Folders$Foci_detection_folder, glob = "*.csv", recurse = TRUE)

focis_all <- map_df(detected_foci_files, \(x) read_csv(x), .id = "focis_detected_filename") |>
    janitor::clean_names() |>
    mutate(
      # condition = fs::path_dir(focis_detected_filename) |> fs::path_dir() |> fs::path_file(),
      # replicate = fs::path_dir(focis_detected_filename) |> fs::path_file(),
      focis_image_name = fs::path_file(focis_detected_filename) |> fs::path_ext_set("tif")
      ) |>
    filter(
      contrast_ch1 >= settings$Analysis_settings$Focis$Good_focis_thresholds$Minimum_contrast,
      snr_ch1      >= settings$Analysis_settings$Focis$Good_focis_thresholds$Minimum_signal_to_noise_ratio,
      quality      >  settings$Analysis_settings$Focis$Good_focis_thresholds$Minimum_quality) 

focis_all |>
    group_by(condition ,replicate, focis_image_name) |>
    group_split() |>
    purrr::map(~sfheaders::sf_point(.x, x = "position_x", y ="position_y", keep =TRUE)) -> l_focis


st_geometry(meshes) <- "mesh_buffered"

meshes |>
  select(
    # condition,
    # replicate,
    focis_image_name, id, mesh_buffered) |>
  group_by(
    # condition,
    # replicate,
    focis_image_name) |>
  group_split() |>
  purrr::map(~select(.x, mesh_buffered, id))-> l_mesh
  
st_geometry(meshes) <- "mesh"

focis_inside <- purrr::map2_dfr(l_focis, l_mesh, ~points_in_mesh(.x,.y)) 

focis_inside_count <- focis_inside |>
  count(id, frame, .drop = FALSE) |>
  complete(
    nesting(id), frame,
    fill = list(n = 0)) 

rm(l_focis, l_mesh)
gc()

}
```



# Diffuse signal analysis

This chunk uses the cytosolic signal (tdTomato) in Bdellovibrio to quantify the growth of Bdellovibrio: the number of pixels inside each Bdellovibrio that are above the threshold (set as 6*MAD, see file 'parameters.yml') is used to calculate the Bdellovibrio area in um2. Non-growing Bdellovibrio are discarded.

```{r}
if (!is.null(settings$Analysis_settings$Folders$Diffuse_folder)) {

l_meshes  <- get_mesh_list(meshes, mesh_buffered, id, diffuse_image_path)
img_paths <- get_image_list(meshes, diffuse_image_path)

extract_image_pixels_from_meshes(img_paths, l_meshes, include_cols = c("id"), parallel = TRUE) -> fluo_intensity

fluo_intensity |>
  group_by(id) |>
  mutate(
    threshold = predator_pix_fct(
      value[frame %in% settings$Analysis_settings$Diffuse$Predator_pixels_threshold$Frames_to_compute],
      coverage_fraction[frame %in% settings$Analysis_settings$Diffuse$Predator_pixels_threshold$Frames_to_compute],
      n =  settings$Analysis_settings$Diffuse$Predator_pixels_threshold$k)) |>
  ungroup() |>
  filter(value >= threshold) -> filtered_pixels

rm(fluo_intensity, l_meshes)
gc()

  
growing_cells <- filtered_pixels |> 
  dplyr::filter(frame > 1) |>
  group_by(id, frame, threshold) |>
  summarize(
    n_pixels = sum(coverage_fraction)) |>
  ungroup() |>
  complete(
    id, frame,
    fill = list(n_pixels = 0)) |>
  arrange(id, frame) |>
  group_by(id) |>
  fill(threshold, .direction = "updown") |>
  mutate(
    is_local_peak = splus2R::peaks(
      n_pixels,
      span =  settings$Analysis_settings$Diffuse$Growth$End_of_growth_peaks_finding_span,
      strict = FALSE,
      endbehavior = 1), ##span is the range the local maximum is searched (now: 12frames up/down = 
    has_peak = TRUE %in% is_local_peak) |> #if there is no peak, the cell is discarded
  dplyr::filter(has_peak) |>
  dplyr::filter(n_pixels > 10) |>
  mutate(
    first_peak = min(frame[which(is_local_peak == TRUE & frame >= settings$Analysis_settings$Diffuse$Growth$Min_first_peak_frame)]),
    # first_peak = find_min(frame, for_which = is_local_peak == TRUE & frame >= settings$Analysis_settings$Diffuse$Growth$Min_first_peak_frame), # to try
    movie_duration = max(frame)) |>
  dplyr::filter(!is.infinite(first_peak)) |>
  dplyr::filter(frame <= first_peak) |>
  mutate(
    bdellovibrio_area    = n_pixels * (settings$Analysis_settings$Microscopy_info$Pixel_to_um_ratio^2),
    perc_movie_growth_duration = max(frame)/movie_duration *100,
    growth_duration_frame = n(),
    perc_growth_duration_under_min_area = length(bdellovibrio_area[bdellovibrio_area < settings$Analysis_settings$Diffuse$Growth$Bdellovibrio_minimum_area_threshold_um])/growth_duration_frame * 100,
    start_size   = first(bdellovibrio_area),
    central_size = central_size_fct(bdellovibrio_area), # From RvR
    end_size     = last(bdellovibrio_area),
    grows        = central_size >= settings$Analysis_settings$Diffuse$Growth$Start_size_threshold * start_size & end_size >= settings$Analysis_settings$Diffuse$Growth$End_size_threshold * start_size) |>
  dplyr::filter(grows) |>
  dplyr::filter(perc_movie_growth_duration >= settings$Analysis_settings$Diffuse$Growth$Percentage_movie_growth_duration_threshold) |>
  dplyr::filter(perc_growth_duration_under_min_area <= settings$Analysis_settings$Diffuse$Growth$Growth_duration_under_min_area_threshold)


rm(filtered_pixels)
gc()

}
```


```{r}

if (exists("popping_frames")) {
saveRDS(popping_frames,
        fs::path(settings$Analysis_settings$Folders$Analysis_folder) / "popping_frames.rds")
}
if (exists("growing_cells")) {
saveRDS(growing_cells,
        fs::path(settings$Analysis_settings$Folders$Analysis_folder) / "growing_cells.rds")
}
if (exists("meshes")) {
saveRDS(meshes,
        fs::path(settings$Analysis_settings$Folders$Analysis_folder) / "meshes_ids.rds")
}
if (exists("focis_inside")) {
saveRDS(focis_inside,
        fs::path(settings$Analysis_settings$Folders$Analysis_folder) / "focis_inside.rds")
}
if (exists("focis_inside_count")) {
saveRDS(focis_inside_count,
        fs::path(settings$Analysis_settings$Folders$Analysis_folder) / "focis_inside_counts.rds")
}

```

