# Downstream analysis & plotting of ParB *B. bacteriovorus* with *E. coli* prey grown in LB

## Library loading

```{r}
library(tidyverse)
library(sf)
```

## Parameters

```{r}
n_bins <- 3
condition_order <- c("WT", "Joe", "William", "Averell") # TODO ? change into the proper strain names ?
save_format <- "csv" # TODO ? Add choice of output and change function with a save_fct<- switch(save_format, ...)
summarise_fcts <- c(
  mean   = \(x) mean(x, na.rm = TRUE),
  sd     = \(x) sd(x, na.rm = TRUE),
  median = \(x) median(x, na.rm = TRUE),
  mad    = \(x) mad(x, na.rm = TRUE),
  CV     = \(x) sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE),
  rCV    = \(x) mad(x, na.rm = TRUE) / median(x, na.rm = TRUE))
```

## Data loading

Load .rds data from the prey size analysis pipeline & also load .rds data containing info on cell to discard after manual checking of outliers cells.

```{r}
analysis_folder <- "E:/Yoann/Citro/" |> fs::path()


if (fs::file_exists(analysis_folder/"meshes_ids.rds")) {
meshes <- readRDS(analysis_folder/"meshes_ids.rds")}
if (fs::file_exists(analysis_folder/"popping_frames.rds")) {
popping_frames <- readRDS(analysis_folder/"popping_frames.rds")}
if (fs::file_exists(analysis_folder/"focis_inside.rds")) {
focis_inside <- readRDS(analysis_folder/"focis_inside.rds")}
if (fs::file_exists(analysis_folder/"focis_inside_counts.rds")) {
focis_inside_count <- readRDS(analysis_folder/"focis_inside_counts.rds")}
if (fs::file_exists(analysis_folder/"phase_intensity_summary.rds")) {
phase_intensity_summary <- readRDS(analysis_folder/"phase_intensity_summary.rds")}
```

## Data merging

Merge data from the outline file (id, bdelloplast outline, condition, replicate, bdelloplast area) with popping point data (id, popping frame/time), diffuse signal (id, bdellovibrio area, frame/time, end of growth) & discard the "bad" cells

```{r}
meshes |> 
  inner_join(popping_frames, by = "id") |>
  inner_join(focis_inside_count, by = "id") |>
  ungroup()  -> ParB_LB_Citrobacter
```

## Data cleaning

```{r}
ParB_LB_Citrobacter |> 
  as_tibble() |>
  filter(
    frame > 1) |>
  group_by(id, condition) |>
  mutate(
    has_early_focis = any(frame %in% 2:3 & n_foci >= 1),
    perc_time_low_focis_number = sum(n_foci %in% 0:1) / n()) |>
  filter(
    !has_early_focis,
    perc_time_low_focis_number <= 0.9) |>
  group_by(id, condition, replicate, bdelloplast_area, popping_time) |>
  summarize(
    max_foci = max(n_foci),
    relative_progenies_per_area = unique(max_foci/bdelloplast_area),
    start_of_first_replication = find_min(time, n_foci >= 1),
    start_of_last_replication = find_min(time, n_foci == max_foci)) |> 
  ungroup() |>
  mutate(
    condition = fct_relevel(condition, condition_order),
    log_bdelloplast_area = log(bdelloplast_area),
    bin_prey_area = santoku::chop_equally(bdelloplast_area, n_bins),
    replication_time = start_of_last_replication - start_of_first_replication,
    difference_popping_time_last_replication = popping_time - start_of_last_replication) -> ParB_LB_Citrobacter
```

## Downstream analysis (can be moved to the pipeline)

```{r}

```

## Descriptive statistics (Summarized data)

Compute the descriptive statistics (mean, SD, CV, median, mad, rCV) of the data across everything, condition, bins of the prey area, number of foci

-   Data \|\> group() \|\> summarize() -\> *uniform way of naming with data name and group* (for plotting the statistics)

*uniform way of naming with data name and group \|\>* pivot_longer() -\> show & save (for easier reading)

### Across all data

```{r}
# TODO make sure everything interesting is summarized
# TODO add sgr to the summary 
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        max_foci,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_LB_Citrobacter_over_all


ParB_LB_Citrobacter_over_all |>
  pivot_longer(
    cols = -n,
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")

```

### Across strains (condition)

```{r}
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(condition) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        max_foci,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_LB_Citrobacter_over_condition


ParB_LB_Citrobacter_over_condition |>
  pivot_longer(
    cols = -c(n, condition),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")
```

### Across replicates (should be important)

```{r}
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(replicate) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        max_foci,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_LB_Citrobacter_over_replicate


ParB_LB_Citrobacter_over_replicate |>
  pivot_longer(
    cols = -c(n, replicate),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")
```

### Across strain & replicates (also important)

```{r}
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(condition, replicate) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        max_foci,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_LB_Citrobacter_over_condition_and_replicate

ParB_LB_Citrobacter_over_condition_and_replicate |>
  pivot_longer(
    cols = -c(n, condition, replicate),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")
```

### Across prey size bins

```{r}
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(bin_prey_area) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        max_foci,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_LB_Citrobacter_over_bins


ParB_LB_Citrobacter_over_bins |>
  pivot_longer(
    cols = -c(n, bin_prey_area),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")
```

### Across prey max foci & condition/bins (for bubble plots)

```{r}
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(max_foci, condition) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_LB_Citrobacter_over_max_foci_and_condition


ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(max_foci, bin_prey_area) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_LB_Citrobacter_over_max_foci_and_bins



```

### Saving

```{r}
# TODO do a loop or a purrr::walk
TdTomato_LB_over_all |> write_csv(file = analysis_folder / "TdTomato_LB_over_all.csv")
```

## Data plotting

### Popping point

```{r}
# TODO Make sure overall plots parts are rougthly in the same order
# Should we add someting like a pch over replicates or another plot with color = replicate ?
# Maybe we should make a list like that (plot_name = plot) maybe nested (not sure) so we can loop / purrr::walk to save

# Popping time/prey area scatter plot
ParB_LB_Citrobacter |>
  ggplot(aes(
    x = bdelloplast_area,
    y = popping_time,
    color = condition)) +
  geom_point() + 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,700)

# Popping time/ln(prey area) scatter plot
ParB_LB_Citrobacter |>
  ggplot(aes(
    x = log_bdelloplast_area,
    y = popping_time)) +
  geom_point(aes(
    color = condition)) + 
  geom_smooth(
    method = "lm",
    color = "black") + 
  theme_classic() + 
  theme(aspect.ratio=1)

# Popping time across conditions
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = condition,
    y = popping_time,
    color = condition)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = popping_time_median ,
    ymin = popping_time_median - popping_time_mad,
    ymax = popping_time_median + popping_time_mad,
    group = condition),
    color = "black",
    data = ParB_LB_Citrobacter_over_condition)+ 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,600)

# Popping time across bins of bdelloplast area
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = bin_prey_area,
    y = popping_time,
    color = bin_prey_area)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = popping_time_median,
    ymin = popping_time_median - popping_time_mad,
    ymax = popping_time_median + popping_time_mad,
    group = bin_prey_area),
    color = "black",
    data = ParB_LB_Citrobacter_over_bins)+ 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,600)
```

### Number of offspring

```{r}
ParB_LB_Citrobacter |>
  ggplot(aes(
    x = bdelloplast_area,
    y = max_foci,
    color = condition)) +
  geom_point() + 
  theme_classic() + 
  theme(aspect.ratio=1)

ParB_LB_Citrobacter |>
  ggplot(aes(
    x = log_bdelloplast_area,
    y = max_foci,
    color = condition)) +
  geom_point() + 
  theme_classic() + 
  theme(aspect.ratio=1)

# Popping time across conditions
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = condition,
    y = max_foci,
    color = condition)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = max_foci_median ,
    ymin = max_foci_median - max_foci_mad,
    ymax = max_foci_median + max_foci_mad,
    group = condition),
    color = "black",
    data = ParB_LB_Citrobacter_over_condition)+ 
  theme_classic() + 
  theme(aspect.ratio=1) 

ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = bin_prey_area,
    y = max_foci,
    color = condition)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = max_foci_median ,
    ymin = max_foci_median - max_foci_mad,
    ymax = max_foci_median + max_foci_mad,
    group = bin_prey_area),
    color = "black",
    data = ParB_LB_Citrobacter_over_bins)+ 
  theme_classic() + 
  theme(aspect.ratio=1) 


ParB_LB_Citrobacter_over_max_foci_and_condition |>
  ggplot(aes(
    x = condition,
    y = max_foci,
    size = n,
    color = condition)) +
  geom_point(alpha=0.5) +
  scale_size(range = c(5, 20)) + 
  theme_classic() + theme(aspect.ratio=1)

ParB_LB_Citrobacter_over_max_foci_and_bins |>
  ggplot(aes(
    x = bin_prey_area,
    y = max_foci,
    size = n,
    color = bin_prey_area)) +
  geom_point(alpha=0.5) +
  scale_size(range = c(5, 20)) + 
  theme_classic() + theme(aspect.ratio=1)
```

### Number of offspring/area

```{r}
ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = condition,
    y = relative_progenies_per_area,
    color = condition)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = relative_progenies_per_area_median ,
    ymin = relative_progenies_per_area_median - relative_progenies_per_area_mad,
    ymax = relative_progenies_per_area_median + relative_progenies_per_area_mad,
    group = condition),
    color = "black",
    data = ParB_LB_Citrobacter_over_condition)+ 
  theme_classic() + 
  theme(aspect.ratio=1) 

ParB_LB_Citrobacter |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = bin_prey_area,
    y = relative_progenies_per_area,
    color = bin_prey_area)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = relative_progenies_per_area_median ,
    ymin = relative_progenies_per_area_median - relative_progenies_per_area_mad,
    ymax = relative_progenies_per_area_median + relative_progenies_per_area_mad,
    group = bin_prey_area),
    color = "black",
    data = ParB_LB_Citrobacter_over_bins)+ 
  theme_classic() + 
  theme(aspect.ratio=1) 
```

### Segregation timing (1^st^ ParB foci - last ParB foci)

```{r}
ParB_LB_Citrobacter |>
ggplot(aes(
    x = start_of_last_replication,
    y = start_of_first_replication)) +
  geom_point(aes(color = condition)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
ggplot(aes(
    x = start_of_last_replication,
    y = start_of_first_replication)) +
  geom_point(aes(color = bin_prey_area)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
  ggplot(aes(
    x = replication_time)) +
  geom_density() + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
  ggplot(aes(
    x = replication_time)) +
  geom_density(aes(color = condition)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
  ggplot(aes(
    x = replication_time)) +
  geom_density(aes(color = bin_prey_area)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0))

```

### Popping point - last ParB foci

```{r}
ParB_LB_Citrobacter |>
ggplot(aes(
    x = popping_time,
    y = start_of_last_replication)) +
  geom_point(aes(color = condition)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
ggplot(aes(
    x = popping_time,
    y = start_of_last_replication)) +
  geom_point(aes(color = bin_prey_area)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
  ggplot(aes(
    x = difference_popping_time_last_replication)) +
  geom_density() + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
  ggplot(aes(
    x = difference_popping_time_last_replication)) +
  geom_density(aes(color = condition)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0))

ParB_LB_Citrobacter |>
  ggplot(aes(
    x = difference_popping_time_last_replication)) +
  geom_density(aes(color = bin_prey_area)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0))

```
