# Downstream analysis & plotting of TdTomato *B. bacteriovorus* with *E. coli* prey grown in LB

## Library loading

```{r}
library(tidyverse)
library(sf)
```

## Data loading

Load .rds data from the prey size analysis pipeline & also load .rds data containing info on cell to discard after manual checking of outliers cells.

```{r}

```

## Data merging

Merge data from the outline file (id, bdelloplast outline, condition, replicate, bdelloplast area) with popping point data (id, popping frame/time), diffuse signal (id, bdellovibrio area, frame/time, end of growth) & discard the "bad" cells

```{r}

```

## Data cleaning

```{r}

```

## Downstream analysis (can be moved to the pipeline)

Compute the specific growth rate (µ) of B. bacteriovorus filaments by fitting a linear model : $ln(area_{prey})(t) = µ \cdot t + ln(area_{prey_{0}})$

```{r}

```

## Descriptive statistics (Summarized data)

Compute the descriptive statistics (mean, SD, CV, median, mad, rCV) of the data across everything, condition, bins of the prey area

-   Data \|\> group() \|\> summarize() -\> *uniform way of naming with data name and group* (for plotting the statistics)

-   *uniform way of naming with data name and group \|\>* pivot_longer() -\> show & save (for easier reading)

```{r}

```

## Data plotting

### Popping point

```{r}

```

### End of growth

```{r}

```

### Popping point - end of growth

```{r}

```

### Individual growth

```{r}

```

### Specific growth rate

```{r}

```
