---
author: Thomas Lamot
title: Visualisation of Bdellovibrio growth dynamics
---

### Prerequisites

This set of code is part of *insert paper title here*. The full set of analysis scripts, templates & data can be found at <https://github.com/giatomo/prey_size_analysis> and uses the output of `Pipeline_final.Rmd`.

To install dependencies run the following chunck:

```{r}
install.packages(setdiff(c('sf', 'tidyverse', 'fs', 'santoku', 'ggbeeswarm'), rownames(installed.packages())))  
```

# Downstream analysis & plotting of TdTomato *B. bacteriovorus* with *E. coli* prey grown in LB

## Library loading

```{r}
library(tidyverse)
library(sf)
```

## Parameters

```{r}
n_bins <- 3
condition_order <- c("WT", "Joe", "William", "Averell") # TODO ? change into the proper strain names ? 
  ##RR I think it's not terribly important, as long as it's understandable
save_format <- "csv" # TODO ? Add choice of output and change function with a save_fct<- switch(save_format, ...)
summarise_fcts <- c(
  mean   = \(x) mean(x, na.rm = TRUE),
  sd     = \(x) sd(x, na.rm = TRUE),
  median = \(x) median(x, na.rm = TRUE),
  mad    = \(x) mad(x, na.rm = TRUE),
  CV     = \(x) sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE),
  rCV    = \(x) mad(x, na.rm = TRUE) / median(x, na.rm = TRUE))
```

## Data loading

Load .rds data from the prey size analysis pipeline & also load .rds data containing info on cell to discard after manual checking of outliers cells. Change `analysis_folder` parameter to correct path to data. Make sure downstream paths are as set below or change paths below accordingly.

```{r}
analysis_folder <- "E:/Yoann/M9_1462/" |> fs::path()

if (fs::file_exists(analysis_folder/"meshes_ids.rds")) {
meshes <- readRDS(analysis_folder/"meshes_ids.rds")}
if (fs::file_exists(analysis_folder/"growing_cells.rds")) {
growing_cells <- readRDS(analysis_folder/"growing_cells.rds")}
if (fs::file_exists(analysis_folder/"popping_frames.rds")) {
popping_frames <- readRDS(analysis_folder/"popping_frames.rds")}
if (fs::file_exists(analysis_folder/"phase_intensity_summary.rds")) {
phase_intensity_summary <- readRDS(analysis_folder/"phase_intensity_summary.rds")}
if (fs::file_exists(analysis_folder/"bad_high_sgr.rds")) {
bad_high_sgr <- readRDS(analysis_folder / "bad_high_sgr.rds")}
if (fs::file_exists(analysis_folder/"bad_early_pop.rds")) {
bad_early_pop <- readRDS(analysis_folder / "bad_late_pop.rds")}
if (fs::file_exists(analysis_folder/"bad_early_pop.rds")) {
bad_late_pop <- readRDS(analysis_folder / "bad_late_pop.rds")}

```

## Data merging

Merge data from the outline file (id, bdelloplast outline, condition, replicate, bdelloplast area) with popping point data (id, popping frame/time), diffuse signal (id, *Bdellovibrio* area, frame/time, end of growth) & discard the "bad" cells.

```{r}
meshes |> 
  # Discard id from manually curated data
  filter(!id %in% c(bad_early, bad_late)) |>
  # Join all the data together 
  inner_join(popping_frames, by = "id") |>
  inner_join(growing_cells, by = "id") |>
  as_tibble() -> TdTomato_M9



TdTomato_M9 |> 
  # Keep only one line of data for each id to create the bins
  # Avoid having and impact of the number of frames when creating *n_bins* bins of the same size 
  group_by(id) |>
  slice_head(n = 1) |>
  ungroup() |>
  mutate(
    bin_prey_area = santoku::chop_equally(bdelloplast_area, n_bins)) |>
  select(id, bin_prey_area) -> TdTomato_M9_bins

```

## Data cleaning & calculation of parameters for plotting

```{r}

TdTomato_M9 |> 
  inner_join(TdTomato_M9_bins, by = "id") |> 
  mutate(
    log_bdellovibrio_area = log(bdellovibrio_area),
    log_bdelloplast_area  = log(bdelloplast_area),
    prey_size_relative_popping_time     = popping_time / bdelloplast_area,
    log_prey_size_relative_popping_time = popping_time / log_bdelloplast_area,
    maturation_time = popping_time - end_of_growth_time,
    condition = fct_relevel(condition, condition_order)) -> TdTomato_M9

# TODO select the most useful infos


```

## Downstream analysis (can be moved to the pipeline)

Compute the specific growth rate (µ) of *B. bacteriovorus* filaments by fitting a linear model : $ln(area_{prey})(t) = µ \cdot t + ln(area_{prey_{0}})$

```{r}
TdTomato_M9 |> 
  group_by(id) |>
  filter(!is.infinite(log_bdellovibrio_area)) |> 
  nest(data = -id) |>
  # Fit lm  and get  the slope
  # TODO Might be useful to keep/add more info (goodness of fit, intercept, etc)
  mutate(
    sgr_min = map_dbl(data, \(data) lm(log_bdellovibrio_area ~ time, data = data)$coefficients[[2]]),
    sgr_h = sgr_min * 60) -> specific_growth_rate

TdTomato_M9 |> inner_join(specific_growth_rate, by = "id") -> TdTomato_M9


# Maybe we should just group this result with the main data and save only once (will also simplifies the descriptive statistics)


```

### Saving

```{r}
# TODO select the most useful infos
TdTomato_M9 |> write_csv(file = analysis_folder / "TdTomato_M9.csv")
```

## Descriptive statistics (Summarized data)

Compute the descriptive statistics (mean, SD, CV, median, mad, rCV) of the data across everything, condition, bins of the prey area

-   Data \|\> group() \|\> summarize() -\> *uniform way of naming with data name and group* (for plotting the statistics)

*uniform way of naming with data name and group \|\>* pivot_longer() -\> show & save (for easier reading)

### Across all data

```{r}
# TODO make sure everything interesting is summarized
# TODO add sgr to the summary 
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        end_of_growth_time,
        start_size,
        central_size,
        end_size,
        difference_popping_point_end_of_growth_min,
        specific_growth_rate
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> TdTomato_M9_over_all


TdTomato_M9_over_all |>
  pivot_longer(
    cols = -n,
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")

```

### Across strains (condition)

```{r}
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(condition) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        end_of_growth_time,
        start_size,
        central_size,
        end_size,
        difference_popping_point_end_of_growth_min,
        specific_growth_rate
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> TdTomato_M9_over_condition


TdTomato_M9_over_condition |>
  pivot_longer(
    cols = -c(n, condition),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")
```

### Across replicates (should be important)

```{r}
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(replicate) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        end_of_growth_time,
        start_size,
        central_size,
        end_size,
        difference_popping_point_end_of_growth_min,
        specific_growth_rate
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> TdTomato_M9_over_replicate


TdTomato_M9_over_replicate |>
  pivot_longer(
    cols -c(n, replicate),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")
```

### Across strain & replicates (also important)

```{r}
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(condition, replicate) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        end_of_growth_time,
        start_size,
        central_size,
        end_size,
        difference_popping_point_end_of_growth_min,
        specific_growth_rate
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> TdTomato_M9_over_condition_and_replicate


TdTomato_M9_over_condition_and_replicate |>
  pivot_longer(
    cols = -c(n, condition, replicate),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")

```

### Across prey size bins

```{r}
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(bin_prey_area) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        end_of_growth_time,
        start_size,
        central_size,
        end_size,
        difference_popping_point_end_of_growth_min,
        specific_growth_rate
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> TdTomato_M9_over_bins


TdTomato_M9_over_bins |>
  pivot_longer(
    cols = -c(n, bin_prey_area),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")
```

### Saving

```{r}
# TODO do a loop or a purrr::walk
TdTomato_M9_over_all |> write_csv(file = analysis_folder / "TdTomato_M9_over_all.csv")

```

## Data plotting

### Popping point

```{r}
# TODO Make sure overall plots parts are rougthly in the same order
# Should we add someting like a pch over replicates or another plot with color = replicate ?
# Maybe we should make a list like that (plot_name = plot) maybe nested (not sure) so we can loop / purrr::walk to save

# Popping time/prey area scatter plot
TdTomato_M9 |>
  ggplot(aes(
    x = bdelloplast_area,
    y = popping_time,
    color = condition)) +
  geom_point() + 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,700)

# Popping time/ln(prey area) scatter plot
TdTomato_M9 |>
  ggplot(aes(
    x = log_bdelloplast_area,
    y = popping_time)) +
  geom_point(aes(
    color = condition)) + 
  geom_smooth(
    method = "lm",
    color = "black") + 
  theme_classic() + 
  theme(aspect.ratio=1)

# Popping time across conditions
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = condition,
    y = popping_time,
    color = condition)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = popping_time_median ,
    ymin = popping_time_median - popping_time_mad,
    ymax = popping_time_median + popping_time_mad,
    group = condition),
    color = "black",
    data = TdTomato_M9_over_condition)+ 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,600)

# Popping time across bins of bdelloplast area
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = bin_prey_area,
    y = popping_time,
    color = bin_prey_area)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = popping_time_median,
    ymin = popping_time_median - popping_time_mad,
    ymax = popping_time_median + popping_time_mad,
    group = bin_prey_area),
    color = "black",
    data = TdTomato_M9_over_bins)+ 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,600)
```

### End of growth

```{r}
TdTomato_M9 |>
  ggplot(aes(
    x = bdelloplast_area,
    y = end_of_growth_time,
    color = condition)) +
  geom_point() + 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,700)

# Popping time/ln(prey area) scatter plot
TdTomato_M9 |>
  ggplot(aes(
    x = log_bdelloplast_area,
    y = end_of_growth_time)) +
  geom_point(aes(
    color = condition)) + 
  geom_smooth(
    method = "lm",
    color = "black") + 
  theme_classic() + 
  theme(aspect.ratio=1)

# Popping time across conditions
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = condition,
    y = end_of_growth_time,
    color = condition)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = end_of_growth_time_median ,
    ymin = end_of_growth_time_median - end_of_growth_time_mad,
    ymax = end_of_growth_time_median + end_of_growth_time_mad,
    group = condition),
    color = "black",
    data = TdTomato_M9_over_condition)+ 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,600)

# Popping time across bins of bdelloplast area
TdTomato_M9 |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = bin_prey_area,
    y = end_of_growth_time,
    color = bin_prey_area)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = end_of_growth_time_median,
    ymin = end_of_growth_time_median - end_of_growth_time_mad,
    ymax = end_of_growth_time_median + end_of_growth_time_mad,
    group = bin_prey_area),
    color = "black",
    data = TdTomato_M9_over_bins)+ 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,600)
```

### Popping point - end of growth

```{r}
# Need to be frames or change KDE method
# Seems to work ok with time on ParB_M9 (could be bc of redundancy in df here ? or idk)
ParB_M9 |>
ggplot(aes(
    x = popping_time,
    y = end_of_growth)) +
  geom_point(aes(color = bin_prey_area)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

ParB_M9 |>
ggplot(aes(
    x = popping_time,
    y = end_of_growth)) +
  geom_point(aes(color = condition)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0))

TdTomato_M9 |>
  ggplot(aes(
      x = difference_popping_point_end_of_growth_min/8)) +
  geom_density(aes(
      color = bin_prey_area)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0, 0.7) + 
  xlim(-200/8, 200/8) 

TdTomato_M9 |>
  ggplot(aes(
      x = difference_popping_point_end_of_growth_min/8)) +
  geom_density(aes(
      color = condition)) + 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0, 0.7) + 
  xlim(-200/8, 200/8) 
```

### Individual growth

```{r}
TdTomato_M9 |>
  ggplot(aes(
    y = bdellovibrio_area,
    x = time)) +
  geom_line(aes(
    color = condition,
    group = id)) + 
  facet_wrap(~condition) + 
  theme_classic() + theme(aspect.ratio=1)

TdTomato_M9 |>
  ggplot(aes(
    y = bdellovibrio_area,
    x = time)) +
  geom_line(aes(
    color = bin_prey_area,
    group = id)) + 
  facet_wrap(~bin_prey_area) + 
  theme_classic() + theme(aspect.ratio=1)

TdTomato_M9 |>
  ggplot(aes(
    y = log_bdellovibrio_area,
    x = time)) +
  geom_line(aes(
    color = condition,
    group = id)) + 
  facet_wrap(~condition) + 
  theme_classic() + theme(aspect.ratio=1)

TdTomato_M9 |>
  ggplot(aes(
    y = log_bdellovibrio_area,
    x = time)) +
  geom_line(aes(
    color = bin_prey_area,
    group = id)) + 
  facet_wrap(~bin_prey_area) + 
  theme_classic() + theme(aspect.ratio=1)

```

### Specific growth rate

```{r}
TdTomato_M9 |>
  ggplot(aes(
    x = bdelloplast_area,
    y = sgr_h)) + 
  geom_point(aes(
    color = condition)) +
  ylim(0,100) +
  theme_classic() + 
  theme(aspect.ratio=1)

TdTomato_M9 |>
  ggplot(aes(
    x = condition,
    y = sgr_h)) + 
  ggbeeswarm::geom_quasirandom(aes(color = condition)) +
  geom_pointrange(aes(
    y = sgr_h_median,
    ymin = sgr_h_median - sgr_h_mad,
    ymax = sgr_h_median + sgr_h_mad),
    color = "black",
    data = TdTomato_M9_over_condition) +
  ylim(0,100) + theme_classic() + theme(aspect.ratio=1)

TdTomato_M9 |>
  ggplot(aes(
    x = bin_prey_area,
    y = sgr_h)) + 
  ggbeeswarm::geom_quasirandom(aes(color = condition)) +
  geom_pointrange(aes(
    y = sgr_h_median,
    ymin = sgr_h_median - sgr_h_mad,
    ymax = sgr_h_median + sgr_h_mad),
    color = "black",
    data = TdTomato_M9_over_bins) +
  ylim(0,100) + theme_classic() + theme(aspect.ratio=1)
```

### Saving

```{r}

```
