# Downstream analysis & plotting of ParB *B. bacteriovorus* with *E. coli* prey grown in LB

## Library loading

```{r}
library(tidyverse)
library(sf)
```

## Parameters

```{r}
n_bins <- 3
condition_order <- c("WT", "Joe", "William", "Averell") # TODO ? change into the proper strain names ?
save_format <- "csv" # TODO ? Add choice of output and change function with a save_fct<- switch(save_format, ...)
summarise_fcts <- c(
  mean   = \(x) mean(x, na.rm = TRUE),
  sd     = \(x) sd(x, na.rm = TRUE),
  median = \(x) median(x, na.rm = TRUE),
  mad    = \(x) mad(x, na.rm = TRUE),
  CV     = \(x) sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE),
  rCV    = \(x) mad(x, na.rm = TRUE) / median(x, na.rm = TRUE))
```


```{r}
analysis_folder <- "E:/Yoann/M9_1462/" |> fs::path()

if (fs::file_exists(analysis_folder/"meshes_ids.rds")) {
meshes <- readRDS(analysis_folder/"meshes_ids.rds")}
if (fs::file_exists(analysis_folder/"growing_cells.rds")) {
growing_cells <- readRDS(analysis_folder/"growing_cells.rds")}
if (fs::file_exists(analysis_folder/"popping_frames.rds")) {
popping_frames <- readRDS(analysis_folder/"popping_frames.rds")}
if (fs::file_exists(analysis_folder/"phase_intensity_summary.rds")) {
phase_intensity_summary <- readRDS(analysis_folder/"phase_intensity_summary.rds")}
if (fs::file_exists(analysis_folder/"bad_high_sgr.rds")) {
bad_high_sgr <- readRDS(analysis_folder / "bad_high_sgr.rds")}
if (fs::file_exists(analysis_folder/"bad_early_pop.rds")) {
bad_early_pop <- readRDS(analysis_folder / "bad_late_pop.rds")}
if (fs::file_exists(analysis_folder/"bad_early_pop.rds")) {
bad_late_pop <- readRDS(analysis_folder / "bad_late_pop.rds")}

meshes |> 
  # Discard id from manually curated data
  filter(!id %in% c(bad_early, bad_late)) |>
  # Join all the data together 
  inner_join(popping_frames, by = "id") |>
  inner_join(growing_cells, by = "id") |>
  as_tibble() -> TdTomato_M9

TdTomato_M9 |> 
  # Keep only one line of data for each id to create the bins
  # Avoid having and impact of the number of frames when creating *n_bins* bins of the same size 
  group_by(id) |>
  slice_head(n = 1) |>
  ungroup() |>
  mutate(
    bin_prey_area = santoku::chop_equally(bdelloplast_area, n_bins)) |>
  select(id, bin_prey_area) -> TdTomato_M9_bins


TdTomato_M9 |> 
  inner_join(TdTomato_M9_bins, by = "id") |> 
  mutate(
    log_bdellovibrio_area = log(bdellovibrio_area),
    log_bdelloplast_area  = log(bdelloplast_area),
    prey_size_relative_popping_time     = popping_time / bdelloplast_area,
    log_prey_size_relative_popping_time = popping_time / log_bdelloplast_area,
    maturation_time = popping_time - end_of_growth_time,
    condition = fct_relevel(condition, condition_order),
    media = "M9") -> TdTomato_M9

analysis_folder <- "E:/Yoann/IGR_all/" |> fs::path() # TODO change path to relative path from github folder or fictive 'data folder'

# Pipeline output
if (fs::file_exists(analysis_folder/"meshes_ids.rds")) {
  meshes <- readRDS(analysis_folder/"meshes_ids.rds")}
if (fs::file_exists(analysis_folder/"growing_cells.rds")) {
  growing_cells <- readRDS(analysis_folder/"growing_cells.rds")}
if (fs::file_exists(analysis_folder/"popping_frames.rds")) {
  popping_frames <- readRDS(analysis_folder/"popping_frames.rds")}
if (fs::file_exists(analysis_folder/"phase_intensity_summary.rds")) {
  phase_intensity_summary <- readRDS(analysis_folder/"phase_intensity_summary.rds")}

# Data issued from manual curation
if (fs::file_exists(analysis_folder/"bad_early.rds")) {
  bad_early <- readRDS(analysis_folder / "bad_early.rds")}
if (fs::file_exists(analysis_folder/"bad_late.rds")) {
  bad_late <- readRDS(analysis_folder / "bad_late.rds")}

meshes |> 
  # Discard id from manually curated data
  filter(!id %in% c(bad_early, bad_late)) |>
  # Join all the data together 
  inner_join(popping_frames, by = "id") |>
  inner_join(growing_cells, by = "id") |>
  as_tibble() -> TdTomato_LB

TdTomato_LB |> 
  # Keep only one line of data for each id to create the bins
  # Avoid having and impact of the number of frames when creating *n_bins* bins of the same size 
  group_by(id) |>
  slice_head(n = 1) |>
  ungroup() |>
  mutate(
    bin_prey_area = santoku::chop_equally(bdelloplast_area, n_bins)) |>
  select(id, bin_prey_area) -> TdTomato_LB_bins


TdTomato_LB |> 
  inner_join(TdTomato_LB_bins, by = "id") |> 
  mutate(
    log_bdellovibrio_area = log(bdellovibrio_area),
    log_bdelloplast_area  = log(bdelloplast_area),
    prey_size_relative_popping_time     = popping_time / bdelloplast_area,
    log_prey_size_relative_popping_time = popping_time / log_bdelloplast_area,
    maturation_time = popping_time - end_of_growth_time,
    condition = fct_relevel(condition, condition_order),
    media = "LB") -> TdTomato_LB


bind_rows(TdTomato_LB, TdTomato_M9) -> TdTomato_all_media

TdTomato_all_media |> 
  group_by(id) |>
  filter(!is.infinite(log_bdellovibrio_area)) |> 
  nest(data = -id) |>
  # Fit lm  and get  the slope
  # TODO Might be useful to keep/add more info (goodness of fit, intercept, etc)
  mutate(
    specific_growth_rate_min = map_dbl(data, \(data) lm(log_bdellovibrio_area ~ time, data = data)$coefficients[[2]]),
    specific_growth_rate_h = specific_growth_rate_min * 60) -> specific_growth_rate

TdTomato_all_media |> inner_join(specific_growth_rate, by = "id") -> TdTomato_all_media

```

```{r}
analysis_folder <- "E:/Yoann/M9_ParB/" |> fs::path()


if (fs::file_exists(analysis_folder/"meshes_ids.rds")) {
meshes <- readRDS(analysis_folder/"meshes_ids.rds")}
if (fs::file_exists(analysis_folder/"popping_frames.rds")) {
popping_frames <- readRDS(analysis_folder/"popping_frames.rds")}
if (fs::file_exists(analysis_folder/"focis_inside.rds")) {
focis_inside <- readRDS(analysis_folder/"focis_inside.rds")}
if (fs::file_exists(analysis_folder/"focis_inside_counts.rds")) {
focis_inside_count <- readRDS(analysis_folder/"focis_inside_counts.rds")}
if (fs::file_exists(analysis_folder/"phase_intensity_summary.rds")) {
phase_intensity_summary <- readRDS(analysis_folder/"phase_intensity_summary.rds")}

meshes |> 
  inner_join(popping_frames, by = "id") |>
  inner_join(focis_inside_count, by = "id") |>
  ungroup() |> 
  as_tibble() |>
  filter(
    frame > 1,
    time <= popping_time) |>
  group_by(id, condition) |>
  mutate(
    has_early_focis = any(frame %in% 2:3 & n_foci >= 1),
    perc_time_low_focis_number = sum(n_foci %in% 0:1) / n()) |>
  filter(
    !has_early_focis,
    perc_time_low_focis_number <= 0.9) |>
  group_by(id, condition, replicate, bdelloplast_area, popping_time) |>
  summarize(
    max_foci = max(n_foci),
    relative_progenies_per_area = unique(max_foci/bdelloplast_area),
    start_of_first_replication = find_min(time, n_foci >= 1),
    start_of_last_replication = find_min(time, n_foci == max_foci)) |> 
  ungroup() |>
  mutate(
    condition = fct_relevel(condition, condition_order),
    log_bdelloplast_area = log(bdelloplast_area),
    bin_prey_area = santoku::chop_equally(bdelloplast_area, n_bins),
    replication_time = start_of_last_replication - start_of_first_replication,
    difference_popping_time_last_replication = popping_time - start_of_last_replication,
    media = "M9") -> ParB_M9

analysis_folder <- "E:/Yoann/PN_combined/" |> fs::path()


if (fs::file_exists(analysis_folder/"meshes_ids.rds")) {
meshes <- readRDS(analysis_folder/"meshes_ids.rds")}
if (fs::file_exists(analysis_folder/"popping_frames.rds")) {
popping_frames <- readRDS(analysis_folder/"popping_frames.rds")}
if (fs::file_exists(analysis_folder/"focis_inside.rds")) {
focis_inside <- readRDS(analysis_folder/"focis_inside.rds")}
if (fs::file_exists(analysis_folder/"focis_inside_counts.rds")) {
focis_inside_count <- readRDS(analysis_folder/"focis_inside_counts.rds")}
if (fs::file_exists(analysis_folder/"phase_intensity_summary.rds")) {
phase_intensity_summary <- readRDS(analysis_folder/"phase_intensity_summary.rds")}


if (fs::file_exists(analysis_folder/"bad_early.rds")) {
bad_early <- readRDS(analysis_folder / "bad_early.rds")}
if (fs::file_exists(analysis_folder/"bad_late.rds")) {
bad_late <- readRDS(analysis_folder / "bad_late.rds")}

if (fs::file_exists(analysis_folder/"bad_big_cell_parB.rds")) {
bad_big_cell_parB <- readRDS(analysis_folder / "bad_big_cell_parB.rds")}
if (fs::file_exists(analysis_folder/"bad_small_cell_parB.rds")) {
bad_small_cell_parB <- readRDS(analysis_folder / "bad_small_cell_parB.rds")}
if (fs::file_exists(analysis_folder/"bad_high_focis_cell_parB.rds")) {
bad_high_focis_cell_parB <- readRDS(analysis_folder / "bad_high_focis_cell_parB.rds")}
if (fs::file_exists(analysis_folder/"bad_low_focis_cell_parB.rds")) {
bad_low_focis_cell_parB <- readRDS(analysis_folder / "bad_low_focis_cell_parB.rds")}
if (fs::file_exists(analysis_folder/"bad_bot_right_quad_cell_parB.rds")) {
bad_bot_right_quad_cell_parB <- readRDS(analysis_folder / "bad_bot_right_quad_cell_parB.rds")}
if (fs::file_exists(analysis_folder/"bad_pop.rds")) {
bad_pop <- readRDS(analysis_folder / "bad_pop.rds")}


corrected_count <- read_csv("C:/Users/distant/Downloads/corrected_foci-data.csv") |> 
  select(id, corrected_count)

meshes |> 
  filter(!id %in% c(bad_big_cell_parB, bad_high_focis_cell_parB, bad_low_focis_cell_parB, bad_small_cell_parB, bad_bot_right_quad_cell_parB)) |>
  inner_join(popping_frames, by = "id") |>
  inner_join(focis_inside_count, by = "id") |>
  left_join(corrected_count, by = "id") |>
  mutate(
    strain = condition,
    condition = case_when(
      condition == "WT_906" ~ "WT",
      condition == "William_906" ~ "William",
      TRUE ~ condition),
    is_bad_pop = if_else(id %in% bad_pop, TRUE, FALSE)) |>
  ungroup() |> 
  as_tibble() |>
  filter(
    frame > 1,
    time <= popping_time) |>
  group_by(id, condition) |>
  mutate(
    has_early_focis = any(frame %in% 2:3 & n_foci >= 1),
    perc_time_low_focis_number = sum(n_foci %in% 0:1) / n()) |>
  filter(
    !has_early_focis,
    perc_time_low_focis_number <= 0.9) |>
  group_by(id, condition, replicate, bdelloplast_area, popping_time) |>
  mutate(
    popping_time = if_else(is.na(corrected_count), popping_time, NA_integer_)) |>
  summarize(
    max_foci = if_else(any(is.na(corrected_count)), max(n_foci), unique(as.integer(corrected_count))),
    relative_progenies_per_area = unique(max_foci/bdelloplast_area),
    start_of_first_replication = find_min(time, n_foci >= 1),
    start_of_last_replication =  if_else(any(n_foci == max_foci), find_min(time, n_foci == max_foci), NA_real_)
    ) |> 
  ungroup() |>
  mutate(
    start_of_last_replication = if_else(is.infinite(start_of_last_replication), NA_real_, start_of_last_replication),
    condition = fct_relevel(condition, condition_order),
    log_bdelloplast_area = log(bdelloplast_area),
    bin_prey_area = santoku::chop_equally(bdelloplast_area, n_bins),
    replication_time = start_of_last_replication - start_of_first_replication,
    difference_popping_time_last_replication = popping_time - start_of_last_replicatio,
    media = "LB") -> ParB_LB



bind_rows(ParB_LB, ParB_M9) -> ParB_all_media
```

## Descriptive statistics (Summarized data)

Compute the descriptive statistics (mean, SD, CV, median, mad, rCV) of the data across everything, condition, bins of the prey area, number of foci

-   Data \|\> group() \|\> summarize() -\> *uniform way of naming with data name and group* (for plotting the statistics)

*uniform way of naming with data name and group \|\>* pivot_longer() -\> show & save (for easier reading)

### Across media

```{r}
# TODO make sure everything interesting is summarized
# TODO add sgr to the summary 
TdTomato_all_media |>
  group_by(id) |> slice_head(n = 1) |> 
  group_by(media) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        end_of_growth_time,
        start_size,
        central_size,
        end_size,
        difference_popping_point_end_of_growth_min,
        specific_growth_rate_h
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> TdTomato_all_media_over_media


TdTomato_all_media |>
  pivot_longer(
    cols = -c(n, media),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")

ParB_all_media |>
  group_by(id) |> slice_head(n = 1) |> ungroup() |>
  group_by(media) |>
  summarise(
    n = n(),
    across(
      c(popping_time,
        max_foci,
        relative_progenies_per_area,
        start_of_first_replication,
        start_of_last_replication,
        replication_time,
        difference_popping_time_last_replication
      ),
      summarise_fcts,
      .names = "{.col}_{.fn}")) -> ParB_all_media_over_media


ParB_all_media_over_media |>
  pivot_longer(
    cols = -c(n, media),
    names_to = c("variable", ".value"),
    names_pattern = "(\\w+)_(\\w+)")

```

### Saving

```{r}
# TODO do a loop or a purrr::walk
TdTomato_LB_over_all |> write_csv(file = analysis_folder / "TdTomato_LB_over_all.csv")
```

## Data plotting

### Popping point

```{r}
# TODO Make sure overall plots parts are rougthly in the same order
# Should we add someting like a pch over replicates or another plot with color = replicate ?
# Maybe we should make a list like that (plot_name = plot) maybe nested (not sure) so we can loop / purrr::walk to save

# Popping time/prey area scatter plot
TdTomato_all_media |>
  ggplot(aes(
    x = bdelloplast_area,
    y = popping_time,
    color = media)) +
  geom_point() + 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,700)

# Popping time/ln(prey area) scatter plot
TdTomato_all_media |>
  ggplot(aes(
    x = log_bdelloplast_area,
    y = popping_time)) +
  geom_point(aes(
    color = media)) + 
  geom_smooth(
    method = "lm",
    color = "black") + 
  theme_classic() + 
  theme(aspect.ratio=1)

# Popping time across conditions
TdTomato_all_media |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = media,
    y = popping_time,
    color = media)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = popping_time_median ,
    ymin = popping_time_median - popping_time_mad,
    ymax = popping_time_median + popping_time_mad,
    group = media),
    color = "black",
    data = TdTomato_all_media_over_media)+ 
  theme_classic() + 
  theme(aspect.ratio=1) + 
  ylim(0,600)

```

### Specific growth rate

```{r}
TdTomato_all_media |>
  ggplot(aes(
    x = bdelloplast_area,
    y = sgr_h)) + 
  geom_point(aes(
    color = media)) +
  ylim(0,100) +
  theme_classic() + 
  theme(aspect.ratio=1)

TdTomato_all_media |>
  ggplot(aes(
    x = media,
    y = sgr_h)) + 
  ggbeeswarm::geom_quasirandom(aes(color = media)) +
  geom_pointrange(aes(
    y = sgr_h_median,
    ymin = sgr_h_median - sgr_h_mad,
    ymax = sgr_h_median + sgr_h_mad),
    color = "black",
    data = TdTomato_all_media_over_media) +
  ylim(0,100) + theme_classic() + theme(aspect.ratio=1)


```

### Number of offspring

```{r}
ParB_all_media |>
  ggplot(aes(
    x = bdelloplast_area,
    y = max_foci,
    color = media)) +
  geom_point() + 
  theme_classic() + 
  theme(aspect.ratio=1)
```

### Number of offspring/area

```{r}
ParB_all_media |>
  group_by(id) |> slice_head(n = 1) |>
  ggplot(aes(
    x = media,
    y = relative_progenies_per_area,
    color = condition)) +
  ggbeeswarm::geom_quasirandom() +
  geom_pointrange(aes(
    y = relative_progenies_per_area_median ,
    ymin = relative_progenies_per_area_median - relative_progenies_per_area_mad,
    ymax = relative_progenies_per_area_median + relative_progenies_per_area_mad,
    group = media),
    color = "black",
    data = ParB_all_media_over_media)+ 
  theme_classic() + 
  theme(aspect.ratio=1) 
```

